\LARGE\singlespacing{\textbf{S1 Appendix} \\ \medskip Spatiotemporal modeling of \emph{Mycobacterium tuberculosis} transmission risks using environmental, clinical, and patient movement data}

### 2024-03-04 doc/paper/manuscript.tex
Updates from Overleaf

% close contacts
Our spatiotemporal model can be viewed as a modified version of the original Wells-Riley transmission model\cite{Riley1978AJE}. It is motivated by previous findings that prolonged close contact is required for transmission\cite{Leung2020NatMed,Brankston2007LancetID,Narasimhan2013PulmonaryMed}. As a result, our model assigns a higher risk of infection to clinical attendees that spend more time in close proximity to infectious individuals. It could thus be used to assess the effectiveness of interventions targeting a reduction in the number of close contacts. Although such assessments only show the potential (modeled) impact of interventions, they can still provide a first estimate about how contact reductions translate into lower transmission risks in crowded indoor settings.

% more complex models and our relation
Despite our modeling efforts, transmission events remain difficult to predict. Their occurrence depends on many other factors such as airflow, the properties of the pathogen, and variation in infectiousness of the source and susceptibility of the host\cite{Wang2021Science}. We discuss these factors in more depth in Supplementary Text~\zref{sec:depth-discussion} in \supp. Importantly, there is large stream of literature considering airflow using computational fluid dynamics (CFD) models to simulate the spatial spread of infectious aerosols in indoor environments, \eg Ref. \cite{Vuorinen2020SafSci,Jung2021InfectChemo,Li2021BuildEnv}. CFD models can also be combined with the Wells-Riley model to estimate the risk of infection\cite{Yan2023BE,Qian2009BE,Li2022SOTTE}, thus are seemingly appealing to our setting as well. However, because they are computationally expensive, CFD models are often only used for experiments, short simulations, and static environments. They also require detailed knowledge about the indoor space, in particular airflow, to model particle transport. By contrast, our setting presents a dynamic environment with frequent patient movements over multiple days in a naturally ventilated space (varying airflow). Moreover, we run our model multiple times in a large number of Monte Carlo simulations to consider uncertainty in all modeling parameters and unmasked infectious TB patients.

>>>>

% close contacts
Our spatiotemporal model can be viewed as a modified version of the original Wells-Riley transmission model\cite{Riley1978AJE}. It is motivated by previous findings that prolonged close contact is required for transmission\cite{Leung2020NatMed,Brankston2007LancetID,Narasimhan2013PulmonaryMed}. As a result, our model assigns a higher risk of infection to clinical attendees that spend more time in close proximity to infectious individuals. It could thus be used to assess the effectiveness of interventions targeting a reduction in the number of close contacts. Although such assessments only show the potential (modeled) impact of interventions, they can still provide a first estimate about how contact reductions translate into lower transmission risks in crowded indoor settings. Nevertheless, we acknowledge that transmission events are generally difficult to predict. Their occurrence depends on many other factors such as airflow, the properties of the pathogen, and variation in infectiousness of the source and susceptibility of the host\cite{Wang2021Science}. We discuss these factors in more depth in Supplementary Text~\zref{sec:depth-discussion} in \supp.

% more complex models and our relation
An important factor that is omitted from our model is airflow. Poor ventilation settings can generate an airflow where infectious particles circulate rather than being dispersed and removed from the indoor space\cite{Li2021BuildEnv}. There is large stream of literature considering airflow using computational fluid dynamics (CFD) models to simulate the spatial spread of infectious aerosols in indoor environments, \eg Ref. \cite{Vuorinen2020SafSci,Jung2021InfectChemo,Li2021BuildEnv}. CFD models can also be combined with the Wells-Riley model to estimate the risk of infection\cite{Yan2023BE,Qian2009BE,Li2022SOTTE}, thus are seemingly appealing to our setting as well. However, because they are computationally expensive, CFD models are often only used for experiments, short simulations, and static environments. They also require detailed knowledge about the indoor space, in particular airflow, to model particle transport. By contrast, our setting presents a dynamic environment with frequent patient movements over multiple days in a naturally ventilated space (varying airflow). Moreover, we run our model multiple times in a large number of Monte Carlo simulations to consider uncertainty in all modeling parameters and unmasked infectious TB patients.

### 2024-03-04 doc/paper/supplements.tex
Updates from Overleaf

% data and other limitations
While we combined multiple data sources to allow spatiotemporal modeling of the generation, spread, and removal of infectious quanta, our data has also limitations. First, clinical attendees could not always be tracked perfectly throughout the clinic using our video sensors. We invested considerable effort to manually link individual tracking IDs probably belonging to the same attendee back together (Supplementary Text~\zref{sec:setting-and-data} in \supp), but some measurement error remains. Second, we linked video tracking and clinical patient data using the times clinical attendees were at the registration and the entry times in the clinical database. False linkages could result when a tracking ID only passed by the registration (without registering) or when multiple IDs were at the registration closely before a clinical entry. Third, to facilitate modeling, we considered the waiting room and corridor as separate, although the door between both rooms was usually open (Supplementary Text~\zref{sec:estimation} in \supp). Fifth, our video sensors could not cover the whole area of the entrance and we did not collect data in the care rooms and other places of the clinic. As a result, the time spent in the clinic and the exposure to infectious quanta may be underestimated for some clinical attendees, although we expect the largest exposure to be in the waiting room, which was fully covered.

>>>>

% data and other limitations
% While we combined multiple data sources to allow spatiotemporal modeling of the generation, spread, and removal of infectious quanta, our data has also limitations. First, clinical attendees could not always be tracked perfectly throughout the clinic using our video sensors. We invested considerable effort to manually link individual tracking IDs probably belonging to the same attendee back together (Supplementary Text~\zref{sec:setting-and-data} in \supp), but some measurement error remains. Second, we linked video tracking and clinical patient data using the times clinical attendees were at the registration and the entry times in the clinical database. False linkages could result when a tracking ID only passed by the registration (without registering) or when multiple IDs were at the registration closely before a clinical entry. Third, to facilitate modeling, we considered the waiting room and corridor as separate, although the door between both rooms was usually open (Supplementary Text~\zref{sec:estimation} in \supp). Fifth, our video sensors could not cover the whole area of the entrance and we did not collect data in the care rooms and other places of the clinic. As a result, the time spent in the clinic and the exposure to infectious quanta may be underestimated for some clinical attendees, although we expect the largest exposure to be in the waiting room, which was fully covered.
